---
title: "Recitation 2: replication"
output: html_document

---
<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In your writing project, you must use the "cleaned" files to replicate tables from a published economics paper. The cleaned files contain a stata file that has two steps:

1. load data
2. run regressions

The regression output (or some subset of the output) will be a column in your table. The tricky part will be to understand the stata code so that you can recreate it in R or excel. This document intends to help you with this task by showing you how to replicate a table from a published economics paper.

You will want to use the R function "felm", whose help page is here:
https://www.rdocumentation.org/packages/lfe/versions/2.8-7/topics/felm


At the end of this document there is a "Tips and hints" section.

# Gonzalez 2012


In this document, I will perform a replication of the paper "The Effect of a Universal Child Benefit on Conceptions,
Abortions, and Early Maternal Labor Supply", by Libertad Gonzalez. I will replicate table 2 of the paper.

The files "Gonzalez_cleaned.zip" are available in the shared drive. They contain three stata files:

- table2_row1.dta
- table2_row2.dta
- tables.do

The first two files contain the data, the do file contains the stata code.

## The .do file

The tables.do file is as follows

```
clear all

use "table2_row1.dta", clear

xi: reg ln post i.post|mc i.post|mc2 i.post|mc3 days if mc>-91 & mc<30, robust
xi: reg ln post i.post|mc i.post|mc2 days if mc>-31 & mc<30, robust
xi: reg ln post i.post|mc i.post|mc2 days if mc>-13 & mc<12, robust
xi: reg ln post i.post|mc days if mc>-10 & mc<9, robust
xi: reg ln post days if mc>-4 & mc<3, robust
xi: reg ln post i.post|mc i.post|mc2 i.post|mc3 days i.month if mc>-91 & mc<30, robust /* 10 years */
xi: reg ln post i.post|mc i.post|mc2 days i.month if mc>-67 & mc<30, robust /* 8 years */
xi: reg ln post i.post|mc i.post|mc2 days i.month if mc>-31 & mc<30, robust /* 5 years */


use "table2_row2.dta", clear

xi: reg log_ive post i.post|m i.post|m2 i.post|m3 days, robust
xi: reg log_ive post i.post|m i.post|m2 days if m>-31, robust
xi: reg log_ive post i.post|m i.post|m2 days if m>-13 & m<12, robust
xi: reg log_ive post i.post|m days if m>-10 & m<9, robust
xi: reg log_ive post days if m>-4 & m<3, robust
xi: reg log_ive post i.post|m i.post|m2 i.post|m3 days i.month, robust /* 10 years */
xi: reg log_ive post i.post|m i.post|m2 days i.month if m>-67, robust /* 8 years */
xi: reg log_ive post i.post|m i.post|m2 days i.month if m>-31, robust /* 5 years */
```

## Row 1, column 1

The first step is to load the data.

```{r echo=T, message=F}
library(tidyverse)
library(haven)
require(stargazer)
require(lfe)
row1.data <- read_dta('table2_row1.dta')
row1.data
```

Now we need to try to understand the first regression. First, let's look at the paper. It "Table 2 reports the results from estimating equations (1) and (2), in eight different
specifications." Equations (1) and (2) are:

\begin{align}
C_m =& \alpha + \gamma_1{m}+\gamma_2(m\cdot {post})+\beta\cdot{post}+\lambda{X}_m+\epsilon_m\\
C_m =& \alpha + \gamma_1{m}+\gamma_2(m\cdot {post})+\beta\cdot{post}+\lambda{X}_m+\sum_{c=2}^{12}\mu_c{month}_m+\epsilon_m\\
\end{align}

The paper describes what each variable is.

Once we work out what the variables mean, we could try regressing it on the first subsample. But first, let's make sure R recognizes that the $post$ variable is a dummy variable ($post$ indicates if the birth took place before or after the cutoff date.)

```{r}
row1.data$post <- factor(row1.data$post)
```

This means R will treat the variable as a dummy variable and not as a number. For the example in class (L7 slide 17), if the "region" variable was coded as $NE=1$, $MW=2$, $S=3$ and $W=4$, then setting ```factor(data$region)``` would ensure R realizes they are dummy variables (one for each region), rather than a number (ie. the regression on slide 17 vs the regression on slide 18).


Now let's do the regression based on the equation above.

```{r results='asis'}
felm(ln ~ 1 + post + days + post*mc, data = row1.data, subset = which(mc>-91 & mc<30)) -> reg1
stargazer(reg1, type='html', header=F, 
          dep.var.labels = "Log number of conceptions per month",
          covariate.labels = c("Post 1 July",
                               "Days in month",
                               "Month of birth for Post=0",
                               "Month of birth for Post=1")
          )
```

This doesn't match the table! (We are expecting 0.0514 on "post 1 July"). So let's look at the stata code:

```
xi: reg ln post i.post|mc i.post|mc2 i.post|mc3 days if mc>-91 & mc<30, robust
```


What does this mean? A good first step is to Google the first word "xi", to see what it means. Based on the help page at https://www.stata.com/manuals/rxi.pdf, it creates interaction terms. So

```
i.post|mc i.post|mc2 i.post|mc3
```

Creates the variables 
\begin{eqnarray}
pre\times{mc} &~~~~~& post\times{mc} &~~~~~& pre\times{mc2} &~~~~~& post\times{mc2} &~~~~~& pre\times{mc3} &~~~~~& post\times{mc3}
\end{eqnarray}
where $pre+post=1$---$pre$ indicating if birth was before the cutoff, $post$ indicating if it was after.

The help page for felm says that we can create interactions by writing
```
post:mc
```
Great! Now lets try this regression:
```{r results='asis'}
felm(ln ~ 1 + post + days + post:mc, data = row1.data, subset = which(mc>-91 & mc<30)) -> reg1
felm(ln ~ 1 + post + post:mc + post:mc2 + post:mc3 + days, data = row1.data, subset = which(mc>-91 & mc<30)) -> reg2
stargazer(list(reg1, reg2), type='html', header=F, 
          dep.var.labels = "Log number of conceptions per month"
          # ,covariate.labels = c("Month of birth",
          #                      "Post 1 July",
          #                      "Days in month",
          #                      "Interaction of month and post",
          #                      "Interaction of month$\\^2$ and post",
          #                      "Interaction of month$\\^3$ and post"
                               # )
          )
```

Looks good! The felm documentation shows that we can suppress the factors by writing the regression as follows:

```{r results='asis'}
felm(ln ~ 1 + post + days | post:mc + post:mc2 + post:mc3, data = row1.data, subset = which(mc>-91 & mc<30)) -> reg
stargazer(reg, 
          se=list(summary(reg, robust=T)$coeff[,2]),
          type='html', header=F, 
          dep.var.labels = "Log number of conceptions per month"
          ,covariate.labels = c("Post 1 July",
                               "Days in month")
          )
```

We have also added heteroskedastic robust SE.

## The rest of Row 1

Now let's do the rest of Row 1:

```
xi: reg ln post i.post|mc i.post|mc2 days if mc>-31 & mc<30, robust
xi: reg ln post i.post|mc i.post|mc2 days if mc>-13 & mc<12, robust
xi: reg ln post i.post|mc days if mc>-10 & mc<9, robust
xi: reg ln post days if mc>-4 & mc<3, robust
xi: reg ln post i.post|mc i.post|mc2 i.post|mc3 days i.month if mc>-91 & mc<30, robust /* 10 years */
xi: reg ln post i.post|mc i.post|mc2 days i.month if mc>-67 & mc<30, robust /* 8 years */
xi: reg ln post i.post|mc i.post|mc2 days i.month if mc>-31 & mc<30, robust /* 5 years */
```

```{r results='asis'}
col1 <- felm(ln ~ 1 + post + days | post:mc + post:mc2 + post:mc3, data = row1.data, subset = which(mc>-91 & mc<30))
col2 <- felm(ln ~ 1 + post + days | post:mc + post:mc2, data = row1.data, subset = which(mc>-31 & mc<30))
col3 <- felm(ln ~ 1 + post + days | post:mc + post:mc2, data = row1.data, subset = which(mc>-13 & mc<12))
col4 <- felm(ln ~ 1 + post + days | post:mc, data = row1.data, subset = which(mc>-10 & mc<9))
col5 <- felm(ln ~ 1 + post + days | 0, data = row1.data, subset = which(mc>-4 & mc<3))
col6 <- felm(ln ~ 1 + post + days | post:mc + post:mc2 + post:mc3 + factor(month), data = row1.data, subset = which(mc>-91 & mc<30))
col7 <- felm(ln ~ 1 + post + days | post:mc + post:mc2 + factor(month), data = row1.data, subset = which(mc>-67 & mc<30))
col8 <- felm(ln ~ 1 + post + days | post:mc + post:mc2 + factor(month), data = row1.data, subset = which(mc>-31 & mc<30))
row1 <- list(col1,col2,col3,col4,col5,col6,col7,col8)

stargazer(row1, 
          se=lapply(row1,function(x)
            summary(x, robust=T)$coeff[,2]),
          type='html', header=F, digits=4,
          dep.var.labels = "Log number of conceptions per month"
          ,covariate.labels = c("Post 1 July",
                               "Days in month")
          )
```

Notice:

- We had to make $month$ a factor variable, since the author says there are "calendar month dummies". If we didn't do this, R would treat "month" as a number (ie. month 10 is bigger than month 5)

## Row 2

Now we have understood the syntax, row 2 is easy
```
xi: reg log_ive post i.post|m i.post|m2 i.post|m3 days, robust
xi: reg log_ive post i.post|m i.post|m2 days if m>-31, robust
xi: reg log_ive post i.post|m i.post|m2 days if m>-13 & m<12, robust
xi: reg log_ive post i.post|m days if m>-10 & m<9, robust
xi: reg log_ive post days if m>-4 & m<3, robust
xi: reg log_ive post i.post|m i.post|m2 i.post|m3 days i.month, robust /* 10 years */
xi: reg log_ive post i.post|m i.post|m2 days i.month if m>-67, robust /* 8 years */
xi: reg log_ive post i.post|m i.post|m2 days i.month if m>-31, robust /* 5 years */
```

The R code is:

```{r results='asis'}
row2.data <- read_dta('table2_row2.dta')
row2.data$post <- factor(row2.data$post)

col1 <- felm(log_ive ~ 1 + post + days | post:m + post:m2 + post:m3, data = row2.data, subset = which(m>-91 & m<30))
col2 <- felm(log_ive ~ 1 + post + days | post:m + post:m2, data = row2.data, subset = which(m>-31 & m<30))
col3 <- felm(log_ive ~ 1 + post + days | post:m + post:m2, data = row2.data, subset = which(m>-13 & m<12))
col4 <- felm(log_ive ~ 1 + post + days | post:m, data = row2.data, subset = which(m>-10 & m<9))
col5 <- felm(log_ive ~ 1 + post + days | 0, data = row2.data, subset = which(m>-4 & m<3))
col6 <- felm(log_ive ~ 1 + post + days | post:m + post:m2 + post:m3 + factor(month), data = row2.data, subset = which(m>-91 & m<30))
col7 <- felm(log_ive ~ 1 + post + days | post:m + post:m2 + factor(month), data = row2.data, subset = which(m>-67 & m<30))
col8 <- felm(log_ive ~ 1 + post + days | post:m + post:m2 + factor(month), data = row2.data, subset = which(m>-31 & m<30))
row2 <- list(col1,col2,col3,col4,col5,col6,col7,col8)

stargazer(row2, 
          se=lapply(row2,function(x)
            summary(x, robust=T)$coeff[,2]),
          type='html', header=F, digits=4,
          dep.var.labels = "Log number of conceptions per month"
          ,covariate.labels = c("Post 1 July",
                               "Days in month")
          )
```


Notice a couple of things:

- Now the "month" variable is called "m" instead of "mc".
- We had to set the dummy variables as factors, just like before.
- There is a slight mismatch between my standard errors and the published standard errors!


# Tips and hints

The main task will be to translate the stata regressions into R. A typical Stata regression will look like this:
```
regress y x1 x2
```
Which means to regress ```y``` on ```(x1,x2)```.

I suggest you use the R function felm to run linear regression. The general syntax for ```felm``` is explained on the felm help page under ```Details```. Type ```?felm``` to get to the help page. The general syntax is:
```{r}`r ''`
felm(y ~ x1 + x2 | f1 + f2 | (x3 + x4 ~ z3 + z4) | c1 + c2, data=my.data)
```
As explained on the help page:

- ```x1``` and ```x2``` are exogenous covariates
- ```f1``` and ```f2``` are categorical variables (called "factors" in ```R```), for which felm will create dummy variables for each level of the factor (for example: ```f1``` could be an industry category (mining, agriculture), felm will create a dummy for each category) (We will cover categorical variables and dummies in the "non-linear regression" lectures.)
- In ```(x3 + x4 ~ z3 + z4)```, ```x3``` and ```x4``` are non-exogenous covariates, and ```z3``` and ```z4``` are instrumental variables (We will cover this in the instrumental variable lectures.)
- ```c1``` and ```c2``` are cluster variables for creating standard errors (We will cover this in the panel lectures.)

The next section translates the regression syntax you will encounter:


## Regression syntax:
The left column is Stata, the right is R

<div class="col2">
```
reg y x1 x2
```
```
reg y x1 if(x2==0 & x3==1)
reg y x1 i.x2
reg y x1 [aweight=x2]
reg y x1 [pweight=x2]
```
```
xi: reg y x1 i.x2*x3
~
```
```
reg y x1 i.x2##c.x3
```
```
areg y x1, absorb(x2)
```
```
ivreg y x1 (x2=x3)
```
```
xtset id_var
xtreg y x1 x2, fe
xtivreg y x1 (x2 = x3), fe
```
```
xtreg y x1 L(0/1).x2
~
~
```
```
reg y x1 x2, robust
~
```
```
reg y x1 x2 x3, cluster(x2 x3)
~
```


<!-- R STARTS HERE -->
```{r}`r ''`
felm(y ~ x1 + x2)
```
```{r}`r ''`
felm(y ~ x1, subset = which(x2==0 & x3==1))
felm(y ~ x1 + factor(x2))
felm(y ~ x1, weight = x2)
felm(y ~ x1, weight = x2)
```
```{r}`r ''`
felm(y ~ x1 + x2:x3)
felm(y ~ x1 | x2:x3)
```
```{r}`r ''`
felm(y ~ x1 + factor(x2):x3)
```
```{r}`r ''`
felm(y ~ x1 | x2)
```
```{r}`r ''`
felm(y ~ x1 | 0 | (x2 ~ x3))
```
```{r}`r ''`
~
felm(y ~ x1 + x2 | id_var) 
felm(y ~ x1 | id_var | (x2 ~ x3))
```
```{r}`r ''`
felm(y ~ x1 + x2 + lag(x2, 1), data=
 pdata.frame(my.data,index=c("ID.var", "time.var")))
# lag is from the "plm" package
```
```{r}`r ''`
reg1 <- felm(y ~ x1 + x2)
summary(reg1, robust = T)
```
```{r}`r ''`
reg1 <- felm(y ~ x1 + x2 + x3 | 0 | 0 | x2 + x3)
summary(reg1, robust = T)
```

</div>

### Variable lists in Stata

- To avoid repitition, Stata defines lists by the following syntax, which can then be called:
```
local NAME1 x1 x2 x3
global NAME2 x4 x5
regress y `NAME1' `NAME2' x6
```
This is equivalent to writing
```
regress y x1 x2 x3 x4 x5 x6
```


### Dummy and categorical variables

- Be sure to make your dummy and categorical variables are factors.
- There are two ways to control for dummy variables using software:
  1. Manually create a dummy variable for each level (eg. MidWest=1 if region=4, South=1 if region=3, \dots), and then include each dummy variable in the regression.
  2. Tell your software the variable is a "factor", and the software will create the dummies for you.
  
It is much simpler to do 2, even if the Stata code takes option 1. If the author chose option 1, the data file will typically still have the factor variable---so you can take option 1, even if the authors take option 2. It will be easier to code if you find the factor variable and include it as $factor(\text{variable.name})$.

### Defining new variables.

In Stata and R:

<div class="col2">

```
gen x1 = x2/2
```
```{r}`r ''`
x1 <- x2/2
```

</div>


### Not equals

Sometimes you will want to subset based on which observations don't have $x1$ equal to 0, for example. This is how it is done in Stata and R:

<div class="col2">

```
x1~=0
```
```{r}`r ''`
x1!=0
```
</div>

### Missing data

In Stata missing data is indicated as follows:

```
x1==.
```

This means the observations that have missing values for the variable ```x1```. In R this would be written ```is.na(x1)```. 

However, keep in mind it may not be necessary to manually exclude missing data, since the ```felm``` function will automatically drop observations with missing variables.

### Dealing with the * syntax in Stata

In Stata, you may see something like this:
```
reg y x1 begin*end x2
```
(Or with either begin or end missing.) This means that Stata will look through the variable names and include all that start with ```begin``` and end with ```end```. There is no equivalently simple way to include all such variables in R, so here are two options. First, you can use R to look up all the variables that match the pattern, and then manually include them in your formula. Here is the required code:
```{r}`r''`
vars.start <- colnames(my.data)[startsWith(colnames(my.data), "begin")]
vars.start_and_end <- vars.start[endsWith(vars.start, "end")]
print(vars.start_and_end)
```
You can then manually add the variable names in ```vars.start_and_end``` to your formula.

Alternately, we can automate the writing of the formula
```{r}`r''`
reg.formula <- paste0("y ~ x1 ", 
                  paste0(vars.start_and_end, collapse="+"),
                  "x2")
felm(as.formula(reg.formula), data=my.data)
```
Option two appears preferable if there are many variables that match ```begin*end```.

### The - syntax in Stata
Similar to ```*```, you may see ```-``` in Stata. For example, ```me1-me6```. This means all six variables from ```me1``` through ```me6```. The two options for handling ```*``` can be used to handle ```-```.


