---
title: "PS2_Submit"
output: pdf_document
date: "2025-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question A: Homogeneous Treatment Effects Model

### 1. Describe the Unit-Level Treatment Effect (ITE)

- **Mathematical Representation**:  
  The individual treatment effect (ITE) is defined as:  
  $$ \text{ITE}_i = Y_i(1) - Y_i(0) $$  
  where \( Y_i(1) \) represents the outcome (spending) for customer \( i \) if they are a member, and \( Y_i(0) \) represents the outcome if they are not.

- **Verbal Representation**:  
  The ITE captures the difference in Walmart spending for an individual customer if they are a Sam's Club Plus member versus if they are not.

### 2. Prove \( Y_i = \alpha + \text{ITE}_iD_i + U_i \)

Using the Rubin Causal Model (RCM), the observed outcome can be expressed as:  
\[
Y_i = Y_i(0)(1 - D_i) + Y_i(1)D_i
\]

Expanding this equation:  
\[
Y_i = Y_i(0) + (Y_i(1) - Y_i(0))D_i
\]

Let:  
- \( \alpha = Y_i(0) \): The baseline outcome (spending) for non-members.  
- \( \text{ITE}_i = Y_i(1) - Y_i(0) \): The individual treatment effect.  

Substituting these definitions gives:  
\[
Y_i = \alpha + \text{ITE}_iD_i + U_i
\]

where \( U_i \) represents the error term, capturing unobserved factors influencing \( Y_i \). Assuming exogeneity, \( \mathbb{E}[U_i] = 0 \).

**Examples of \( U_i \)**:  
\( U_i \) could include factors such as:
- Household size.
- Income level.
- Shopping preferences.

### 3. Compute the Bias of \( \hat{\beta}_n \) for \( \beta \)

**Definition of Bias**:  
\[
\text{Bias}_{\theta}(\hat{\theta}_n) = \mathbb{E}[\hat{\theta}_n] - \theta
\]

Using the independence assumption \( \{Y_i(1), Y_i(0)\} \perp D_i \):  
\[
\text{Bias} = \left[\mathbb{E}[Y_i(1) \mid D_i = 1] - \mathbb{E}[Y_i(0) \mid D_i = 0]\right] - \left[\mathbb{E}[Y_i(1)] - \mathbb{E}[Y_i(0)]\right]
\]

This simplifies to:  
\[
\text{Bias} = \left[\mathbb{E}[Y_i(1)] - \mathbb{E}[Y_i(0)]\right] - \left[\mathbb{E}[Y_i(1)] - \mathbb{E}[Y_i(0)]\right] = 0
\]

Thus, under the independence assumption, the bias is zero.

### 4. Bias with Simplified Expressions

Given \( \hat{\beta} = \mathbb{E}[Y_i \mid D_i = 1] - \mathbb{E}[Y_i \mid D_i = 0] \), the bias can be rewritten as:  
\[
\text{Bias} = \left[\mathbb{E}[Y_i \mid D_i = 1] - \mathbb{E}[Y_i \mid D_i = 0]\right] - \left[\mathbb{E}[Y_i(1)] - \mathbb{E}[Y_i(0)]\right]
\]

Under the overlap condition \( 0 < \Pr(D_i = 1 \mid X) < 1 \), this expression ensures no bias because the treatment and control groups are comparable within strata of \( X \).

---

## Question B: Regression Model for ATU

### 1. Specify \( \alpha \) and \( U_i \)

To model the Average Treatment Effect for the Untreated (ATU):  
\[
Y_i = \alpha + \text{ATU} \cdot D_i + U_i
\]

Here:  
- \( \alpha = \mathbb{E}[Y_i(0)] \): The average outcome for non-treated individuals.  
- \( U_i \): Captures deviations from the average untreated outcome, including unobserved factors.

### 2. Exogeneity Condition

The exogeneity condition requires that \( \mathbb{E}[U_i \mid D_i] = 0 \), ensuring that the treatment assignment does not depend on unobserved confounders.

---

## Question C: Conditional Random Assignment (CRA)

### 1. Correlation of Variables with Treatment

- **Number of Walmart visits (likely correlated)**: Customers who visit Walmart frequently may perceive higher benefits from membership.  
- **Amazon spending (uncorrelated)**: Membership at Walmart likely does not influence spending on Amazon.  
- **Number of children (likely correlated)**: Families with children may visit Walmart more often, leading to higher membership uptake.

### 2. Checking the CRA Assumption

To check CRA:  
- Stratify the data by tenure levels.  
- Perform \( t \)-tests on pre-determined variables (e.g., number of visits, number of children) to assess balance across treatment groups.  

### 3. Identification of CATE

CATE is identified when:  
1. Conditional independence: \( \{Y(1), Y(0)\} \perp D \mid X \).  
2. Overlap: \( 0 < \Pr(D = 1 \mid X = x) < 1 \).

### 4. Lack of Overlap

If \( \Pr(D = 1 \mid X = x) = 0 \) or \( 1 \), treatment and control groups are not comparable, making CATE unidentifiable.

### 5. Estimating ATE

**Code Example**:
```{r}
estimate_ate <- function(data) {
  tenure_groups <- unique(data$tenureD)
  
  cate_estimates <- sapply(tenure_groups, function(x) {
    group_data <- subset(data, tenureD == x)
    mean(group_data$spend[group_data$scp == 1]) - 
    mean(group_data$spend[group_data$scp == 0])
  })
  
  emp_dist <- table(data$tenureD) / nrow(data)
  ate <- sum(cate_estimates * emp_dist)
  
  return(ate)
}
```

## Question D

```{r}
library(dplyr)
library(ggplot2)
walmart_data <- read.csv("SCP_data.csv", stringsAsFactors = FALSE)
# Part 1: tenure distirbtuion
tenure_plot <- ggplot(walmart_data, aes(x = tenure)) + 
  geom_histogram(bins = 10, alpha = 0.7, fill = "blue") + 
  ggtitle("Customer Tenure Distribution") +
  xlab("Tenure") +
  ylab("Number of Customers")
print(tenure_plot)

# Given the distribution being skewed, 
# we would not want to say that the indepdence assumption is gone
```

```{r}
# Part 2: discrete assignment
walmart_data <- walmart_data %>% mutate(
  tenureD = case_when(
    tenure <= 1 ~ 1,
    tenure > 5 ~ 3,
    TRUE ~ 2
  )
)
# Discrete size tenure should be independence, 
# and we could create a more uniform distribution based on so
```

```{r}
# Part 3: Conditional Average Treatment Effect Estimation
cate <- lm(spend ~ scp + factor(tenureD), data = walmart_data)
cate

```


```{r}
# Part 4: Estimator Comparison

# dsicrete partition CATE
d_partition <- walmart_data %>%
  group_by(tenureD) %>%
  summarize(
    mean_spend_scp_1 = mean(spend[scp == 1]),
    mean_spend_scp_0 = mean(spend[scp == 0]),
    partition_ate = mean_spend_scp_1 - mean_spend_scp_0
  )
    
d_partition


# difference in expectant values (assumes homogeneity)
exp_diff <- mean(walmart_data$spend[walmart_data$scp == 1], na.rm = TRUE) - 
  mean(walmart_data$spend[walmart_data$scp == 0], na.rm = TRUE)
exp_diff


# The partition/ tenure D is more preferred because we assume heterogeneity effect; 
# mean difference appraoch may overestiamte the effect of the memership in spending 
```

```{r}
# Part 5: Comparison
true_ate <- mean(walmart_data$spend1 - walmart_data$spend0)

d_partition_value <- mean(d_partition$partition_ate) # Aggregate to a single value
comparison <- data.frame(
  method = c("Discrete CATE", "Difference in Expected Values", "Real ATE"),
  value_estimate = c(d_partition_value, exp_diff, true_ate)
)
print(comparison)




```